Verifying Concurrent Search Structure Templates for C Programs using VST

### 2023-04-20 concurrency/CONCUR 2023/paper.tex
combine two figure into one

\subsection{Search Structure Templates}
\label{templates}
The C implementation of a node in our template style is:
\begin{lstlisting}[language = C, backgroundcolor=\color{white}, basicstyle=\ttfamily\footnotesize]
typedef struct node_t {node *t; lock_t *lock;} node_t;
\end{lstlisting}
\than{I changed \lstinline{tree_t} into \lstinline{node_t}, and \lstinline{tree} into \lstinline{node}, since concurrent search structure will not depend on particular data structures.}

@@

\begin{figure}[ht]
	\begin{subfigure}[t]{0.45\textwidth}
		\lstinputlisting[language=caml, style=myStyle]{lock_traverse.ml} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in an ML-like language \cite{krishna2019compositional}} 
			%\wm{Shouldn't it be lock and then unlock?}} \than{I copied exactly the same from thesis of Krishna. However, in the their of proofs, they had lock before unlock like you said. Do we need to changed to fit with our context and their code?} 
		\label{traverse_lock_a}	
	\end{subfigure}\qquad
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{lock_traverse.c} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in C}
		\label{traverse_lock_b}
	\end{subfigure}
	\caption{The \lstinline{traverse} method of the lock-coupling template algorithms}
	\label{traverse_lock}
\end{figure}

@@

The \texttt{inRange} specification (Figure \ref{fig:inRange_giveup}) simply necessitates that \texttt{inRange(pn, k)}  yields a boolean value, \texttt{res}. When \texttt{res} is \texttt{True}, it signifies that \texttt{k} is within the range of \texttt{(min, max)} associated with the node \texttt{n}. 
Conversely, \texttt{res} is \texttt{False}, it indicates that \texttt{k} is outside the boundaries of \texttt{(min, max)}.

As previously mentioned, the \texttt{insertOp} specification (Figure \ref{fig:insertOp_giveup}) exhibits slight differences from that of the lock-coupling template, primarily due to the incorporation of two extra fields. As a result, node \texttt{n} needs point-to predicates for these new fields. Moreover, the current node's \texttt{l} and \texttt{r} point to two new empty child nodes with associated locks and ranges: \texttt{(NULL, lk1, (min, k))} and \texttt{(NULL, lk2, (k, max))}.  In the context of binary search trees, a node inherits its range from its parent based on the parent's key; thus, if node \texttt{n} has range \texttt{(min, max)} and key \texttt{k}, its left child will have the range \texttt{(min, k)} and its right child the range \texttt{(k, max)}.

\begin{figure}
	\centering
\begin{mathpar}
	{\color{blue}
		\left\{ 
		\begin{array}{c}
			\texttt{pn}\mapsto\texttt{n} \ \ast \texttt{n}\mapsto\texttt{(t, lock, min, max)} 
		\end{array}
		\right\}
	}
	\\ 
	\texttt{inRange(pn, k)} 
	\\
	{\color{blue}
		\left\{\exists \  \texttt{res.}
		\begin{array}{c}
			 \ \texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{(t, lock, min, max)} \  \\ 
			 \ \ast  \ \texttt{(if res then } \texttt{(min < k} \land \texttt{k < max)} \land \texttt{(k} \leq \texttt{min} \lor \texttt{k} \geq \texttt{max))}
		\end{array}
		\right\}
	}
\end{mathpar}
\caption{Specification of \texttt{inRange} for the give-up template}
\label{fig:inRange_giveup}
\end{figure}

\begin{figure}
	\centering
	\begin{mathpar}
	{\color{blue}
		\left\{ 
		\begin{array}{c}
			\texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{t} \  \ast \ \texttt{n}\mapsto\texttt{min} \ \ast \ \texttt{n}\mapsto\texttt{max}
		\end{array}
		\right\}
	}
	\\
	\texttt{insertOp(pn, k, v)}
	\\
	{\color{blue}
	\left\{\exists \ \texttt{t' l r lk1 lk2}. 
	\begin{array}{c}
		\ \texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{t'} \ \ast \ \texttt{t'}\mapsto \texttt{(k, v, l, r)} \\ 
		\ \ast \ \texttt{l} \mapsto \texttt{(NULL, lk1, (min, k))} \ \ast \ \texttt{r} \mapsto \texttt{(NULL, lk2, (k, max))}
	\end{array}
	\right\}
	}
\end{mathpar}
\caption{Specification of \texttt{insertOp} for the give-up template}
\label{fig:insertOp_giveup}
\end{figure}

>>>>

\subsection{Search Structure Templates}
\label{templates}
The C implementation of a node in our template style is:
\begin{lstlisting}[language = C, backgroundcolor=\color{white}, basicstyle=\ttfamily\footnotesize]
typedef struct node_t {node *t; lock_t *lock;} node_t;
\end{lstlisting}

@@

\begin{figure}[ht]
	\begin{subfigure}[t]{0.45\textwidth}
		\lstinputlisting[language=caml, style=myStyle]{lock_traverse.ml} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in an ML-like language \cite{krishna2019compositional}} 
		\label{traverse_lock_a}	
	\end{subfigure}\qquad
	\begin{subfigure}[t]{0.48\textwidth}
		\lstinputlisting[language=C, style=myStyle]{lock_traverse.c} 
		\caption{The \lstinline{traverse} method of the lock-coupling template algorithm written in C}
		\label{traverse_lock_b}
	\end{subfigure}
	\caption{The \lstinline{traverse} method of the lock-coupling template algorithms}
	\label{traverse_lock}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{mathpar}
		% inRange for give-up
		{\color{blue}
			\left\{ 
			\begin{array}{c}
				\texttt{pn}\mapsto\texttt{n} \ \ast \texttt{n}\mapsto\texttt{(t, lock, min, max)} 
			\end{array}
			\right\}
		}
		\\ 
		\texttt{inRange(pn, k)} 
		\\
		{\color{blue}
			\left\{\exists \  \texttt{res.}
			\begin{array}{c}
				\ \texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{(t, lock, min, max)} \  \\ 
				\ \ast  \ \texttt{(if res then } \texttt{(min < k} \land \texttt{k < max)} \land \texttt{(k} \leq \texttt{min} \lor \texttt{k} \geq \texttt{max))}
			\end{array}
			\right\}
		}
		\\
		\\
		%insertOp for give-up
		{\color{blue}
			\left\{ 
			\begin{array}{c}
				\texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{t} \  \ast \ \texttt{n}\mapsto\texttt{min} \ \ast \ \texttt{n}\mapsto\texttt{max}
			\end{array}
			\right\}
		}
		\\
		\texttt{insertOp(pn, k, v)}
		\\
		{\color{blue}
			\left\{\exists \ \texttt{t' l r lk1 lk2}. 
			\begin{array}{c}
				\ \texttt{pn}\mapsto\texttt{n} \ \ast \ \texttt{n}\mapsto\texttt{t'} \ \ast \ \texttt{t'}\mapsto \texttt{(k, v, l, r)} \\ 
				\ \ast \ \texttt{l} \mapsto \texttt{(NULL, lk1, (min, k))} \ \ast \ \texttt{r} \mapsto \texttt{(NULL, lk2, (k, max))}
			\end{array}
			\right\}
		}
	\end{mathpar}
	\caption{Specifications of \texttt{inRange} and \texttt{insertOp} for the give-up template}
	\label{fig:inRange_insertOp_giveup}
\end{figure}

The \texttt{inRange} specification (Figure \ref{fig:inRange_insertOp_giveup}) simply necessitates that \texttt{inRange(pn, k)}  yields a boolean value, \texttt{res}. When \texttt{res} is \texttt{True}, it signifies that \texttt{k} is within the range of \texttt{(min, max)} associated with the node \texttt{n}. 
Conversely, \texttt{res} is \texttt{False}, it indicates that \texttt{k} is outside the boundaries of \texttt{(min, max)}.

As previously mentioned, the \texttt{insertOp} specification (Figure \ref{fig:inRange_insertOp_giveup}) exhibits slight differences from that of the lock-coupling template, primarily due to the incorporation of two extra fields. As a result, node \texttt{n} needs point-to predicates for these new fields. Moreover, the current node's \texttt{l} and \texttt{r} point to two new empty child nodes with associated locks and ranges: \texttt{(NULL, lk1, (min, k))} and \texttt{(NULL, lk2, (k, max))}.  In the context of binary search trees, a node inherits its range from its parent based on the parent's key; thus, if node \texttt{n} has range \texttt{(min, max)} and key \texttt{k}, its left child will have the range \texttt{(min, k)} and its right child the range \texttt{(k, max)}.
